Map{R}educe in {MPI} for Large-scale Graph Algorithms

### 2011-02-23 paper/graph.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@814 3fc257bb-8550-0410-81b3-a5c103da83f7

The algorithm in Figure~\ref{fig:pr2} performs these iterations, using
linear-algebra concepts implemented in a MapReduce framework, with
data stored in four MapReduce objects: $A$, $MT$, $x$, and $y$.  $A$
stores the matrix; $MT$ stores the row index of ``empty'' rows of $A$; 
and $x$ and $y$ are
vectors of length $|V|$.  The nonzeros of matrix $A$ are initialized
as described above; the PageRank vector $x$ is initialized uniformly.
A MapReduce object $MT$ contains the indices of all-zero rows of $A$,
corresponding to vertices with no outlinks.  The PageRank algorithm is
made more efficient by invoking the {\it aggregate()} operation once
on each of these MapReduce objects before the PageRank iterations
begin.  Since the objects are all keyed by vertex, pre-aggregating
them moves all KV pairs with a given key to the same processor.  Many
of the subsequent PageRank operations then become local operations,
replacing a {\it collate()} (which is equivalent to an {\it
aggregate()} followed by {\it convert()}) with simply a {\it
convert()}.  In Figure~\ref{fig:pr2}, steps where {\it convert()} is
substituted for {\it collate()} are marked with an asterisk.  This
strategy of pre-aggregating to improve data locality is useful in many
MapReduce algorithms, but assumes that identical keys in different
MapReduce objects can be mapped to the same processor.

@@

A connected component of a graph is a set of vertices where all pairs
of vertices in the set are connected by a path of edges.  A sparse
graph may contain many such components.  Figure~\ref{fig:cc} outlines
a MapReduce algorithm that labels each vertex in a graph with a
component ID.  All vertices in the same component are labeled with
the same ID, which is the ID of some vertex in the component.  We
assume an input graph $E$ of undirected edges $(V_i,V_j)$.  This
exposition follows the connected-component algorithm presented
in~\cite{Cohen09}, with the addition of logic that load-balances data
across processors, which is important when one or a few giant
components exist in the graph.

@@

The algorithm begins (before the iteration loop) by assigning each
vertex to its own component or ``zone,'' so that $Z_i$ = $V_i$ and
storing this assignment in MapReduce object $V$.  Each iteration grows
the zones, one layer of neighbors at a time.  As zones collide due to
shared edges, a winner is chosen (the smaller zone ID), and vertices
in the losing zone are reassigned to the winning zone.  When the
iterations complete, each zone has become a fully connected component
and the MR object $V$ holds the zone assignment for every vertex.  The
algorithm thus finds all connected components in the graph
simultaneously.  The number of iterations required depends on the
largest diameter of any component in the graph.

>>>>

The algorithm in Figure~\ref{fig:pr2} performs these iterations, using
linear-algebra concepts implemented in a MapReduce framework.  Related
algorithms, each with their own optimizations, are discussed in
\cite{Kang09,Lin10}.  In this version data is stored in four MapReduce
objects: $A$, $MT$, $x$, and $y$.  $A$ stores the matrix; $MT$ stores
the row index of ``empty'' rows of $A$; and $x$ and $y$ are vectors of
length $|V|$.  The nonzeros of matrix $A$ are initialized as described
above; the PageRank vector $x$ is initialized uniformly.  A MapReduce
object $MT$ contains the indices of all-zero rows of $A$,
corresponding to vertices with no outlinks.  The PageRank algorithm is
made more efficient by invoking the {\it aggregate()} operation once
on each of these MapReduce objects before the PageRank iterations
begin.  Since the objects are all keyed by vertex, pre-aggregating
them moves all KV pairs with a given key to the same processor.  Many
of the subsequent PageRank operations then become local operations,
replacing a {\it collate()} (which is equivalent to an {\it
  aggregate()} followed by {\it convert()}) with simply a {\it
  convert()}.  In Figure~\ref{fig:pr2}, steps where {\it convert()} is
substituted for {\it collate()} are marked with an asterisk.  This
strategy of pre-aggregating to improve data locality is useful in many
MapReduce algorithms; it assumes that identical keys in different
MapReduce objects can be mapped to the same processor.

@@

A connected component (CC) of a graph is a set of vertices where all
pairs of vertices in the set are connected by a path of edges.  A
sparse graph may contain many such components.  Figure~\ref{fig:cc}
outlines a MapReduce algorithm that labels each vertex in a graph with
a component ID.  All vertices in the same component are labeled with
the same ID, which is the ID of some vertex in the component.  We
assume an input graph $E$ of undirected edges $(V_i,V_j)$.  Kang
et.~al.~\cite{Kang09} propose a MapReduce CC algorithm based on a
matrix-vector multiplication kernel.  This exposition follows the CC
algorithm presented in~\cite{Cohen09}, with the addition of logic that
load-balances data across processors, which is important when one or a
few giant components exist in the graph.

@@

The algorithm begins (before the iteration loop) by assigning each
vertex to its own component or ``zone,'' so that $Z_i$ = $V_i$ and
storing this assignment in MapReduce object $V$.  Each iteration grows
the zones, one layer of neighbors at a time.  As zones collide due to
shared edges, a winner is chosen (the smaller zone ID), and vertices
in the losing zone are reassigned to the winning zone.  When the
iterations complete, each zone has become a fully connected component
and the MR object $V$ holds the zone assignment for every vertex.  The
algorithm thus finds all CCs in the graph simultaneously.  The number
of iterations required depends on the largest diameter of any
component in the graph.

### 2011-02-23 paper/intro.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@814 3fc257bb-8550-0410-81b3-a5c103da83f7

More recently, MapReduce formulations of traditional number-crunching
kinds of scientific computational tasks have been described, such as
post-processing analysis of simulation data~\cite{Tu08}, graph
algorithmics~\cite{Cohen09}, and linear algebra
operations~\cite{Ekanayake09}.  The paper by Tu et al.~\cite{Tu08} was
particularly insightful to us, because it described how MapReduce
could be implemented on top of the ubiquitous distributed-memory
message-passing interface (MPI), and how the intermediate data-shuffle
operation is conceptually identical to the familiar MPI\_Alltoall
operation.  Their implementation of MapReduce was within a Python
wrapper to simplify the writing of user programs.  The paper motivated
us to develop our own C++ library built on top of MPI for use in graph
analytics, which we initially released as open-source software in
mid-2009.  We have since worked to optimize several of the library's
underlying algorithms and to enable its operation in out-of-core mode
for larger data sets.  These algorithmic improvements are described in
this paper and are part of the current downloadable version; see
Section~\ref{sec:thanks} for details.

>>>>

More recently, MapReduce formulations of traditional number-crunching
kinds of scientific computational tasks have been described, such as
post-processing analysis of simulation data~\cite{Tu08}, graph
algorithmics~\cite{Cohen09,Kang09,Lin10}, and linear algebra
operations~\cite{Ekanayake09}.  The paper by Tu et
al.~\cite{Tu08} was particularly insightful to us, because it
described how MapReduce could be implemented on top of the ubiquitous
distributed-memory message-passing interface (MPI), and how the
intermediate data-shuffle operation is conceptually identical to the
familiar MPI\_Alltoall operation.  Their implementation of MapReduce
was within a Python wrapper to simplify the writing of user programs.
The paper motivated us to develop our own C++ library built on top of
MPI for use in graph analytics, which we initially released as
open-source software in mid-2009.  We have since worked to optimize
several of the library's underlying algorithms and to enable its
operation in out-of-core mode for larger data sets.  These algorithmic
improvements are described in this paper and are part of the current
downloadable version; see Section~\ref{sec:thanks} for details.

### 2011-02-23 paper/results.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@814 3fc257bb-8550-0410-81b3-a5c103da83f7

To give a rough sense of the effort involved to implement the various
MapReduce algorithms described in this section, we list the line
counts of the C++ codes used to perform the benchmarks.  Each code
makes calls to the MR-MPI library and implements the algorithm in the
indicated figure.  The line counts exclude comments and blank lines
which typically add about 30\% to the total.  They also exclude the
routines used to generate the R-MAT matrix input (except in the case
of the two R-MAT algorithms themselves) or process it into
upper-triangular form, since that is common to several of the
benchmarks.

\begin{table}
\begin{center}
\begin{tabular}{|l|c|c|c|c|c|c|c|c|}
\hline
Algorithm & R-MAT & R-MAT++ & PageRank & Tri & CC & MIS & SSSP & SSSP++ \\
\hline
Lines of code & 208 & 216 & 603 & 367 & 453 & 393 & 394 & 448 \\
Figure & \ref{fig:rmat} & \ref{fig:rmat2} & \ref{fig:pr2} & \ref{fig:tri} & \ref{fig:cc} & \ref{fig:luby} & \ref{fig:sssp} & \ref{fig:sssp2} \\
\hline
\end{tabular}
\caption{Lines of code to implement each of the graph
  algorithms in the associated figures.}
\label{t:rmat}
\end{center}
\end{table}

>>>>

To give a rough sense of the effort involved to implement the various
MapReduce algorithms described in this section, we list the line
counts of the C++ codes used to perform the benchmarks in
Table~\ref{t:linecount}.  Each code makes calls to the MR-MPI library
and implements the algorithm in the indicated figure.  The line counts
exclude comments and blank lines which typically add about 30\% to the
total.  They also exclude the routines used to generate the R-MAT
matrix input (except in the case of the two R-MAT algorithms
themselves) or process it into upper-triangular form, since that is
common to several of the benchmarks.

\begin{table}
\begin{center}
\begin{tabular}{|l|c|c|c|c|c|c|c|c|}
\hline
Algorithm & R-MAT & R-MAT++ & PageRank & Tri & CC & MIS & SSSP & SSSP++ \\
\hline
Lines of code & 208 & 216 & 603 & 367 & 453 & 393 & 394 & 448 \\
Figure & \ref{fig:rmat} & \ref{fig:rmat2} & \ref{fig:pr2} & \ref{fig:tri} & \ref{fig:cc} & \ref{fig:luby} & \ref{fig:sssp} & \ref{fig:sssp2} \\
\hline
\end{tabular}
\caption{Lines of code to implement each of the graph
  algorithms in the associated figures.}
\label{t:linecount}
\end{center}
\end{table}
